---
title: "Prediction Assignment Writeup"
author: "Damethri George"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(rpart)
library(randomForest)
library(gbm)
library(e1071)
library(ggplot2)
```

# Executive Summary

This report aims to predict the manner in which participants performed barbell lifts using data collected from accelerometers on various body parts. We use cross-validation and Random Forests to achieve high predictive accuracy and apply the final model to 20 test cases.

# Load and Prepare Data

```{r load-data}
# Download training and test sets if not present
training_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testing_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

train_file <- "pml-training.csv"
test_file <- "pml-testing.csv"

if (!file.exists(train_file)) download.file(training_url, destfile=train_file, method="curl")
if (!file.exists(test_file)) download.file(testing_url, destfile=test_file, method="curl")

training_raw <- read.csv(train_file, na.strings=c("NA", "", "#DIV/0!"))
testing_raw <- read.csv(test_file, na.strings=c("NA", "", "#DIV/0!"))
```

# Data Cleaning

```{r clean-data}
training_raw <- training_raw[, colSums(is.na(training_raw)) == 0]
training_raw <- training_raw[, -(1:7)] # remove irrelevant columns
training_raw$classe <- as.factor(training_raw$classe)

set.seed(1234)
inTrain <- createDataPartition(training_raw$classe, p=0.7, list=FALSE)
training <- training_raw[inTrain,]
testing <- training_raw[-inTrain,]
```

# Model Training

```{r model-training}
set.seed(1234)
control <- trainControl(method="cv", number=5, verboseIter=FALSE)
model_rf <- train(classe ~ ., data=training, method="rf", trControl=control)
model_rf
```

# Model Evaluation

```{r model-evaluation}
predictions <- predict(model_rf, testing)
confusionMatrix(predictions, testing$classe)
```

# Prediction on Test Set

```{r test-prediction}
final_predictions <- predict(model_rf, testing_raw)
final_predictions
```
